---
title: "Test Code"
author: "Rachel Fellman"
date: "2023-12-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(maps)
```

read in data
```{r}
landslide<- read_csv("Global_Landslide_Catalog_Export.csv")
```
# Things to fix 
1) order of factor 

```{r}
land.na$landslide_size <- as.factor(land.na$landslide_size)
land.na$landslide_size <- factor(land.na$landslide_size, levels = c("catastrophic", "very_large", "large", "medium", "small", "unknown"))
```

```{r}
levels(land.na$landslide_size)
```


# data exploration



create map of data

can change color to fatality_count, injury_count, or admin_division_population, need to change name with color
```{r}
#set up global data points
world<- map_data("world")

#create base map layer
a<- ggplot()
#add latitude and longitude information for the map
a.plot<- a + geom_polygon(data = world, aes(x=long, y= lat, group = group), color = "darkgrey", fill = "grey") +
#add the points of landslides
   geom_point(data = landslide, aes(x = longitude, y = latitude, colour = color) ) +
#add a color gradient for the magnitude of the earthquakes
   scale_colour_gradient(low = "green", high = "red", name = "Population")+
#add labels.
  labs(x = "Longitude", y = "Latitude", title = "World Map of Landslides") +
   theme(title = element_text(size=10))
# display plot
a.plot

```


graph of fatality count by size, country name, trigger, or category
```{r}
#create base plot with landslide size on x axis
b <- ggplot(data = landslide, aes(x = as.factor(landslide_size)))
#create bar plot with user input for the fill
b + geom_bar(aes(fill = as.factor(landslide_trigger)))
```

```{r}
g<- ggplot()
      g+ geom_bar(data = land.na, aes(x = landslide_size, fill = as.factor(landslide_setting))) +
      #add labels
        scale_fill_discrete(name = "Setting")
      labs(title = "Bar Graph of Landslide Size")
```

create states maps
```{r}
#set up state data points
state<- map_data("state")

#create base map layer
d<- ggplot()
#add latitude and longitude information for the map
d.plot<- d + geom_polygon(data = state, aes(x=long, y= lat, group = group), color = "darkgrey", fill = "grey") +
#add the points of the earthquakes in California
   geom_point(data = eq.california, aes(x = longitude, y = latitude, colour = magnitude))  +
#add a color gradient for the magnitude of the earthquakes
   scale_colour_gradient(low = "lightblue", high = "darkblue", name = "Magnitude")+
#adjust the coordinates
  coord_cartesian(xlim = c(-127,-110), ylim = c(30,45)) +
#add labels.
  labs(x = "Longitude", y = "Latitude", title = "Plot of Earthquake Locations Colored by Magnitude in a 800 km Radius of the Center of California")
```

```{r}
state<- map_data("state")

#create base map layer
d<- ggplot()
#add latitude and longitude information for the map
d + geom_polygon(data = state, aes(x=long, y= lat, group = group), color = "darkgrey", fill = "grey")
```

remove NA valuse for fatality count.
```{r}
land.us<- landslide %>% 
  filter(country_code == "US") %>% 
  drop_na(fatality_count)
```
 filter some stuff
```{r}
land.na <- landslide %>% 
  drop_na(fatality_count) %>% 
  select("landslide_category","landslide_trigger","landslide_size","landslide_setting","fatality_count","injury_count","country_name","admin_division_population","gazeteer_closest_point", "gazeteer_distance", "longitude", "latitude")
```

scatter of fatality count vs population
```{r}
g <- ggplot()
g + geom_jitter(data = land.na, aes(x = fatality_count, y = admin_division_population, color = landslide_size))
```

numerical summaries

group by option
```{r}
land.na %>% 
      select("fatality_count", "landslide_size", "country_name") %>% 
      filter(country_name == "United States") %>% 
      group_by(landslide_category) %>% 
      summarize(mean = mean(fatality_count), max = max(fatality_count), min = min(fatality_count), sd = sd(fatality_count))
```

factors
```{r}
#make country name a factor
land.na$country_name <- as.factor(land.na$country_name)
#make landslide category a factor
land.na$landslide_category <-as.factor(land.na$landslide_category)
#make landslide trigger a factor
land.na$landslide_trigger <-as.factor(land.na$landslide_trigger)
#make landslide setting a factor
land.na$landslide_setting <-as.factor(land.na$landslide_setting)
```

# modeling
```{r}
names(land.na[,-5])
```

```{r}
 train1 <- reactiveValues({
    # create training index
    train.index<- createDataPartition(land.na$fatality_count, p = (input$test/100), list = FALSE)
    #create train and test sets
    landslide.train <- land.na[train.index, ]
    landslide.test <- land.na[-train.index, ]
    
    #fit mlr model using train function from caret package
    fit.mlr <- train(fatality_count ~ !!sym(input$mlr.vars), data = landslide.train,
                     #select method
                     method = "lm",
                     #preprocess data
                     preProcess = c("center","scale"),
                     #do cross validation
                     trControl = trainControl(method = "cv", number = 5))

  })

  
  output$mlr.train <- renderUI({
    x<-(train1$fit.mlr)
    print(x)
  })
```

```{r}
land.train.sub <- land.nomiss
```


  #use reactive to fit models
  models <- reactive ({
    #call data from train.test
    data <- train.test()
    
```{r}
    #fit mlr model
    fit.mlr <- train(fatality_count ~ landslide_setting + country_name, data = land.nomiss,
                     #select method
                     method = "lm",
                     #preprocess data
                     preProcess = c("center","scale"),
                     #do cross validation
                     trControl = trainControl(method = "cv", number = 5))

```

```{r}
fit.rfor <- train(fatality_count ~ landslide_trigger + country_name , data = land.nomiss,
                  #select method
                  method = "rf",
                  #do cross validation
                 trControl = trainControl(method = "cv", number = 2),
                 #add tuning parameter
                 tuneGrid = data.frame(mtry = 1:9))
```

```{r}
imp <- varImp(fit.rfor, scale = FALSE)
```

```{r}
plot(imp, top = 20)
```

```{r}
    pred.mlr <- predict(fit.mlr, newdata = land.nomiss)
h<-   postResample(pred.mlr, obs = land.nomiss$fatality_count)

```

```{r}
predict(fit.mlr, newdata = data.frame(landslide_category = "landslide", landslide_trigger = "unknown", landslide_setting = "unknown", injury_count = 0, country_name = "United States", admin_division_population = 20000, longitude = -100, latitude = 45))
```


```{r}
print(fit.rfor)
```

    
```{r}
land.nomiss<- land.na[complete.cases(land.na), ]
```
    
```{r}
paste(names(land.nomiss), collapse = "+")
```
    
```{r}
assign(names(land.nomiss$landslide_category),land.nomiss$landslide_category)
```

```{r}
x<- syms(names(land.pred))
land.pred <- land.nomiss[,-5]
```

```{r}
```


```{r}
length(x)
```


```{r}
for (i in 1:length(x)){
  x[[i]] = land.pred[,i]
}
```

```{r}
y<- data.frame(x)
```

